---
title: "BBets Project"
author: "Jaime Duran"
date: "2/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The question is how much money is won/lost on a bet that at least two people have the same b'day in a class of 35.  Instructor wagers 2:1 

Instuctor has done this 15 times,  
1. what is his expected win/loss amount,  
2. how much would the instructor have to wager to make the bet fair 
3. how many students would have to be in the class for a 2:1 bet to be fair?   
 
 
```{r}
days = 1:365
students = 29
classes = 15
runs = students*classes
payout = 2

saveArray=vector('numeric')

for(k in 1:1000){
  bank = 0
  for(i in 1:runs){
    sampleClass = sample(days, students, replace = TRUE)

    repeats = length(unique(sampleClass))
    if (repeats == students) bank = bank - payout
    else bank = bank + 1
  }
saveArray[k]= bank
}

```
 
histogram of bank amounts.   

```{r}
hist(saveArray)
```

Answer question #1

1. The bank is expected to lose 33 times (bank -33), while the professor would win 28 times, because the value for 'repeats' is (repeats 28L). 

```{r}
print('average amt:')
print(mean(saveArray))
```

Answer question # 2

2. To make the wager fair, the professor should wager between 2 to 4 dollars. For the sake of consistency, I ran the simulation where payout is 2 dollars.

Answer question # 3

3. At 35 students, the 2:1 payout ratio was averaging 231.693 on the histogram. The class size that got the best results was 29 students. At a class size of 28 students, the 2 to 1 wager averaged out at -17.016, at 29 students that same payout ratio brought the average size to 17.451. 
